---
title: "Rmarkdown overview"
author: "John Thompson"
date: "29th March 2023"
output:
  xaringan::moon_reader:
    self-contained: true
    css: ["metropolis", "style.css"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Markup Languages

<br>
- A Markup Language inserts formatting instructions into plain text e.g. to produce bold, italics, insert an image etc.   

<br>

- Commonly used markup languages include **html** and **latex**   
<br>

- html and latex can do everything but are complex    
---
# Markdown

<br>
- Markdown is a simple markup language first released in 2004    
  - it only offers very basic formatting  
  - it is very easy to learn   
<br>
- Markdown files have the **.md** extension  
<br>
- Markdown files can be rendered into an html document, pdf, Word document, slide presentation, website, blog, book, thesis, research article   
<br>
- see `RStudio - Help - Markdown Quick Reference`  
---
# YAML

<br>
- YAML (Yet Another Markup Language) is used to create a header for a markdown file  
<br>
- the header controls the format of the final document, e.g. pdf or html, whether or not to add a table of contents, choice of theme etc.
---
# Markdown Example

```{r eval=FALSE, echo = TRUE}
---
title: "Markdown"
author: "John Thompson"
date: "29th March 2023"
output: html_document
---

## Introduction

Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word
documents. For more details see the RStudio menu `Help - Markdown Quick
Reference`.   

When you click the **preview** button in RStudio, the formatted document 
is produced.  
```
---
# Rmarkdown

<br>
- Rmarkdown is an R package that extends markdown by allowing blocks of R Code to be placed within a markdown file  
<br>
- Prior to rendering, the R code is executed and the results are knitted into the markdown file  
<br>
- The blocks of R code are called **chunks**  
<br>
- Rmarkdown files have the **.rmd** extension.  
<br>
- see https://rmarkdown.rstudio.com/ 

---
# Rmarkdown Example

````{r eval=FALSE, echo = TRUE}
---
title: "Rmarkdown"
author: "John Thompson"
date: "29th March 2023"
output: html_document
---

## Introduction

Rmarkdown allows the addition of R code to a markdown file. The chunk below
summarises the cars data frame and adds the results to the document. 
**echo = FALSE** tells rmarkdown to add the results but not the R code.  

```{r echo = FALSE}
summary(cars)
```

When you click the **knit** button in RStudio, the R Code is executed, the
results are inserted and then the formatted document is produced.  
````
