---
title: "Course Organisation"
author: "John Thompson"
date: "9th Febuary 2023"
output:
  xaringan::moon_reader:
    self-contained: true
    css: ["metropolis", "style.css"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
xaringanExtra::use_panelset()
```


# Course Content

<br>
## Original request

-   Teach **git** and **rmarkdown**

<br>
## My Plan

-   Teach **Git**, **GitHub** and **rmarkdown** 
-   Emphasise practical skills that you can use every day  
-   My main aim is **attitudinal change**: the course is only a success, if you come away believing that you have to change the way that you work

---
# I asked ChatGPT

<br>
```{r eval=FALSE, echo=TRUE}
Hi, I have to teach a course on git and github aimed at data analysts and
biostatisticians. What are the key points that I should make to motivate
them to use git and github?
```
---
# ChatGPT Replied

```{r eval=FALSE, echo=TRUE}
Great question! Git and GitHub are powerful tools for version control and
collaboration that can greatly benefit data analysts and biostatisticians. 
Here are some key points that you can use to motivate your students to use Git 
and GitHub:

Version control: Git allows users to keep track of changes made to files
and revert to previous versions if necessary. This can be extremely
useful when working with large datasets or codebases that are constantly
being updated.

Collaboration: GitHub provides a platform for collaborating on projects
with other team members, both locally and remotely. It allows multiple
users to work on the same files at the same time, making it easy to share
code and data and to track changes made by others.

Code reviews: GitHub allows for code reviews, where team members can
review each other's code and suggest changes or improvements. This can
help ensure that code is well-documented, follows best practices, and is
free of errors.
```
---
# More advice from ChatGPT

```{r eval=FALSE, echo=TRUE}
Open source community: GitHub is also home to a large open-source
community where users can find code and contribute to projects. By
contributing to open-source projects, students can build their skills and
network with other professionals in their field.

Portfolio building: By using Git and GitHub to manage their code and
data, students can create a professional portfolio that showcases their
skills and expertise. This can be a valuable asset when applying for jobs
or internships.

Automation: Git and GitHub can be integrated with other tools to automate
tasks such as running tests, building documentation, and deploying code.
This can save time and reduce the risk of errors.

Overall, Git and GitHub are essential tools for data analysts and
biostatisticians who want to work efficiently, collaborate effectively,
and build their skills and portfolio. By mastering these tools, students
can become more productive and successful in their careers.
```
---
# Why Context is Important
<br>
- ChatGPT makes some good points, but I don't think that I'd change my way of working based on such abstract ideas.  
<br>
-   Learning the commands of rmarkdown, git and github is easy .. use Google to look them up when you need them    
<br>
-   Efficiently integrating these skills into your day to day work is more challenging and takes experience  
<br>
-   Raises the more general question, **what is the best way to organise a data analysis project?**

---

# Course Organisation

<br>
-   A brief overview of Git, GitHub and rmarkdown in the context of the best way to organise a data analysis project  
<br>
-   After coffee; I'll create a data analysis project using best practice    
<br>
-   The demonstration project is called CAGE (Classification After Gene Expression)  
<br>
-   The project will use Git, GitHub and rmarkdown  
<br>
-   Your exercise is to use the same best practice on a similar project called RAGE (Regression After Gene Expression)  

---
# It is a Masterclass

-   I will assume that you already know  
    -   basic statistical methods  
    -   how to use R and RStudio  
    -   about microarrays and gene expression  
<br>

-   <p style="color:red">

    Stop me if I assume too much

    </p>

<br>
-   I will avoid teaching things that you can look up on the internet  
<br>
-   Less on **how** you do it and more on **why** and **when**
