---
title: "Setting Up a New Project"
subtitle: "my approach"
author: "John Thompson"
date: "29th March 2023"
output:
  xaringan::moon_reader:
    self-contained: true
    css: ["metropolis", "style.css"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
xaringanExtra::use_panelset()
```


# Choose a Project Name

<br>
- The name should be short, meaningful, memorable eg CAGE  
<br>
- You can waste hours thinking of a good acronym  
<br>
- The same name will be used for the Git and GitHub repos where you will save your work    
<br>
- Create a local folder with this name e.g. C:/Projects/CAGE

---

# Folder Structure

Create the project subfolders. I use this structure for all of my projects  
```{r eval=FALSE, echo = TRUE}
C:/Projects/CAGE ---- code  
                 |  
                 ---- data    ---- archive
                 |            |
                 |            ---- cache
                 |            |
                 |            ---- rawData
                 |  
                 ---- docs    ---- diary
                 |            |
                 |            ---- pdfs
                 |  
                 ---- reports  
                 |  
                 ---- temp 
```


---
# RStudio Project

- Create a new RStudio project using either    
   - `File - New Project`
   - Project button located to the top right - `New Project`  
<br>
- Select `Existing Directory` and browse to the project folder  
<br>
- RStudio creates a file, e.g. CAGE.Rproj, to save information about your project  
<br>
- This RStudio project is now active  

---
# Tell RStudio about Git

If this is the first time that you have used Git then you may need to help RStudio locate it

## Global Options

- Go to global options using the menus `Tools - Global Options`  
<br>
- Select the tab for version control named `Git/SVN`  
<br>
- RStudio may already be aware of Git, if not, 
  - click the `enable version control` box  
  - browse to the Git executable, perhaps `C:/Program Files/Git/bin/git.exe`  
<br>
- Making RStudio aware of the location of Git does not mean that Git is used for every project  

---
# Add Git to this Project

## Project Options

- Use menus `Tools - Project Options` and select the `Git/SVN` tab  
<br>
- The version control window will show `none` meaning that version control is currently off  
<br>
- Change it to `Git`  
<br>
- RStudio will need to restart so that this change can take effect  
<br>
- RStudio will use .Rproj to remember that Git is being used for this project.  

---
# Once Git is Activated

- Git creates a hidden folder called **.git** located within the project folder .. this will contain your archived files   
<br>
- Git creates a text file **.gitignore** that lists files that you do NOT want to archive. Edit it as necessary.  
    - *.html means don't archive any html files  
    - temp/ means don't archive anything from the temp folder  

```{r eval=FALSE, echo=TRUE}
.Rhistory
.RData
*.Rproj
*.html
data/
temp/
docs/
```

---
# Open a Diary

Use either a paper diary or an electronic one  

I keep my diary electronically in a text file called `C:/Projects/RAGE/docs/diary/diary.txt`

This is how I might start the file. Each time I work on the project, I add a dated comment.  
```{r eval=FALSE, echo=TRUE}
Project  CAGE  

Classification After Gene Expression

Aim: investigation of principal component analysis for feature
extraction from gene expression data when the ultimate goal is
disease diagnosis.

17 Feb 2023 
Project folder created with a git archive mirrored on github.
19th Feb 2023
First version of read_data.R to download and read the raw expression data.
```

---
# Create a README

<br>

- GitHub is used to make your work public  
- Readers need to know what the project is about  
- Describe the project in a markdown file **README.md**  
- Start simple and add to the README as the project develops  

<br>

### Examples of README files

* https://github.com/MadryLab/dataset-replication-analysis/blob/master/README.md  

* https://github.com/Lorenzovazq/Cardiovascular-disease-analysis---R

---
# Things to Put in a README

**Simple is good. Don't get carried away. Be Selective**

* The **title** of the project  

* A description of the project, particularly the **aim**  

* The start date and whether the project is still **active**  

* A list of active **collaborators**. An **acknowledgement** of past contributions

* The **progress** to date. What conclusions you have reached so far.  

* A list of **resources** used by the project (GitHub projects don't automatically use R). Perhaps a list of R packages that you rely on with their version numbers. The operating system that you use. 

* The source of the **data** and whether it is public or private.  

---
# More things to put is a README

* **Instructions** on how to run your code. Do some files need to be run before others, or is there a make file that runs everything?

* Key **references**    
  - Only include essential references , preferably with links. 
  - Always include references to publications that stemmed from the project.  

* **Pictures** and logos that help explain the project or which give the project an identity. Remember to add ALT-TEXT for people who are using a screen reader. Careful of copyright.

* GitHub shields, **badges** and icons display information in a very concise way and are popular. Search "github badges" to find lots of resources and tips.

[![Windows](https://svgshare.com/i/ZhY.svg)](https://svgshare.com/i/ZhY.svg)  

---
# LICENSE

- The aim of openness is to make others aware of your work and to share it with them  
- There are two dangers  
  - Someone will use your code without acknowledging you    
  - Someone will use your code, something will go wrong and they will blame you  

- Protect yourself, by releasing your work under license.

- Create a file **LICENSE.md**  

- There are many licences that you could use https://docs.github.com/en/repositories/managing-your-repositorys-settings-and-features/customizing-your-repository/licensing-a-repository 

- I use the MIT license, but you should have a group policy    
---
# MIT LICENSE

```{r eval=FALSE, echo=TRUE}
Copyright <YEAR> <COPYRIGHT HOLDER>

Permission is hereby granted, free of charge, to any person obtaining 
a copy of this software and associated documentation files (the
“Software”), to deal in the Software without restriction, including
without limitation the rights to use, copy, modify, merge, publish,
distribute, sublicense, and/or sell copies of the Software, and to
permit persons to whom the Software is furnished to do so, subject to
the following conditions:
  
The above copyright notice and this permission notice shall be included
in all copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED “AS IS”, WITHOUT WARRANTY OF ANY KIND, EXPRESS
OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF
MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT.
IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY
CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT,
TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE
SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.
```

---
# Make your first Commit

- You have added or edited .gitignore, README.md and LICENSE.md

- Click the `git` tab in RStudio   
  - RStudio lists the files that have changed   
  - Click the boxes next to them, in order to **stage** them  
  - Status `A` means that they are Added to the list of files that will be committed  
- Click the `commit` button    
- Enter a commit message e.g. `Initial Commit`  
- Click the `commit` button, just below the message box  
- Your local Git repo has been updated  

---
# Create a GitHub Repo

<br>

 - Sign in to GitHub  
 - Click the `+` icon in the top right hand corner  
 - Choose `New Repository` and the New Repository Window Opens  
 - Enter the repo name. **It must be the project name**  
 - Skip the invitation to create a README and a LICENCE. You already have them  
 - Go straight to the bottom of the window and click `Create repository`
 - GitHub tells you what to do under the heading `push an existing repository`

---
# Set up the Link

- Back in RStudio with your project active  
- Open a Terminal and type the commands suggested by GitHub. Alternatively you can cut and paste them  
```{r eval=FALSE, echo=TRUE}
git remote add origin https://github.com/thompson575/CAGE.git
git branch -M main
git push -u origin main
```

* `git remote` tells local Git to add a connection to the GitHub repo calling it `origin`  
* `git branch` puts you on the `main` branch of your local repo (-M means Move)  
* `git push` sends the local repo to `origin` from `main`  

- Return to your GitHub account
- .gitignore  README.md and LICENSE.md should have been synced  
- Pushing is usual quite quick. Try page refresh  
- Git will remember the link, so you will not need to set it again  

---
# Do Some Data Analysis

<br>
- You are ready to download the data and start coding     
<br>
- Remember to commit regularly  
---
# Code Structure

<br>
- Separate scripts for each stage in the analysis  
  - read_data.R, pca.R, feature.R, aegis1.R  
<br>
- Each script will save its results in the cache  
<br>
- There will be an rmarkdown report on each stage  
<br>
- RawData => Script1 =>  Cache  => Report1    
- Cache   => Script2  => Cache  => Report2  
<br>
- Scripts produce no other outputs  
<br>
- Repeated tasks are performed by R functions  
  - functions for calculation (used in scripts)  
  - functions for display (used in reports)  
